Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-08-13 Paper.Rmd
adding intro basic

*Goes in the last paragraph of the intro:*
We address the following questions. 1) Do datasets in ecology and related fields have non-Gausian copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? 2) If so, what are some possible causes of non-Normal copula structure? And 3) what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?

# Material and methods

## Datasets

@@

## Theoretical models

Question 2 from the Introduction was explored using simple population models.

@@

# Results

@@

# Discussion

>>>>

Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in the
finance and risk management. For many years, measuring and modeling dependencies has focused on the
Pearson correlation because of its ease in use. But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two random variables. Therefore, it is better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation. However, using copula-based approach as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number. The flexibility of copulas allows us to model more complex dependence structures by isolating its marginal from its joint distribution structures, with the capability in modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry. Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution. For example, in the upper panel of Fig. \ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, were shown which became dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we moved along the row. Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$). To implement this, we need to select candidate bivariate copulas from many existing parametric copula families. A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from. Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.

Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016] or a short period study of intersite dependence of rainfall data [@Serinaldi2008], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data. So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both variable are rare synchronously. If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them. Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.

Basically, we want to address the following questions. 1) Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? 2) If so, what are some possible causes of non-Normal copula structure? And 3) what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Material&Methods}, we described briefly about each datasets we used to answer these questions and provided a detailed descriptions of implemented methods. In the \nameref{Results} section, we summarized our finidings for each datasets. Next two sections are focused on possible causes and consequences of our analysis. Finally, we concluded in the \nameref{Discussion} section.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption{Pedagogical figure showing BB1 copula with varying tail-dep difference but of same Kendall's tau along each row and with same tail-dep difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

# Material and methods \label{Material&Methods}

## Datasets
<!-- each dataset's description is in suppmat file-->

@@

<!--Theoretical models
Question 2 from the Introduction was explored using simple population models.-->

@@

# Results \label{Results}

@@

# Discussion \label{Discussion}

### 2018-08-13 SupportingInformation.Rmd
adding intro basic

for(i in desLTmUT_vector){
  # print(i)
  destau<-0.5
  call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
  obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
  c<-BiCopSim(N=500,obj)
  plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
  mtext(paste0("tau = ",round(obj$tau,5),", LTmUT = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
}

for(i in desLTmUT_vector){
 # print(i)
  destau<-0.8
 call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
 obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
 c<-BiCopSim(N=500,obj)
 plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
 mtext(paste0("tau = ",round(obj$tau,5),", LTmUT = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
}

>>>>

for(i in desLTmUT_vector){
  # print(i)
  destau<-0.5
  call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
  obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
  c<-BiCopSim(N=500,obj)
  plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
  mtext(paste0("tau = ",round(obj$tau,5),", ( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
  legend("topleft",paste0("UT=",round(obj$taildep$upper,5)),bty="n",cex=1.5,text.col="magenta")
  legend("bottomright",paste0("LT=",round(obj$taildep$lower,5)),bty="n",cex=1.5,text.col="magenta")
}

for(i in desLTmUT_vector){
 # print(i)
  destau<-0.8
 call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
 obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
 c<-BiCopSim(N=500,obj)
 plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
 mtext(paste0("tau = ",round(obj$tau,5),", ( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
 legend("topleft",paste0("UT=",round(obj$taildep$upper,5)),bty="n",cex=1.5,text.col="magenta")
 legend("bottomright",paste0("LT=",round(obj$taildep$lower,5)),bty="n",cex=1.5,text.col="magenta")
}
